A numerical study of variational data assimilation for the wave equation in heterogeneous media

### 2024-12-09 tex/main.tex
add paragraph in intro and expand description of numexp

\noindent In this article, we introduce a discontinuous in time finite element method to solve the data assimilation problem \eqref{eq:waveEquation}-\eqref{eq:IF-conditions}+\eqref{eq:dataMatch} in the heterogeneous case. In particular, we want to investigate numerically if the following assumption is justified.

@@

@@

\section{Numerical experiments}
\noindent In this section, we present numerical experiments carried out with the proposed method. The examples are implemented using the \texttt{FEniCSx} library \cite{BarattaEtal2023,BasixJoss,ScroggsEtal2022,AlnaesEtal2014}. 
~\jp{We need to prepare some reproduction material.}
\subsection{Examples in one space dimension}
\noindent We consider the unit interval $\Omega \coloneqq [0,1] \subset \mathbb{R}$. For a refinement level $L \in \mathbb{N}$, we partition the domain into $2^{L+1}$ uniform elements such that $h = 1/2^{L+1}$. The time step is chosen as $\Delta t = h$.

\subsubsection{Simple Example}\label{sec:numex:1D:simple}
\noindent We define the subdomains $\Omega_1 = (0,0.5)$ and $\Omega_2 = (0.5,1.0)$ such that $\Omega = \overline{\Omega}_1 \cup \overline{\Omega}_2$. Then, we fix the wavespeed $c_2 := c \vert_{\Omega_2} = 1$ and consider different values for $c_1 := c \vert_{\Omega_1}$. As a first example, we consider the following exact solution \cite{MHI08}
\begin{equation}\label{eq:1D:exact:simple}
    u(x,t) := \begin{cases}
        \cos(w_1 c_1 t) \cos(w_1(x-0.5)), & x \in \Omega_1, \\
        \cos(w_2 c_2 t) \cos(w_2(x-0.5)), & x \in \Omega_2,
    \end{cases}
\end{equation}
where we set $w_1 = 3 \pi$ and $w_2 = w_1 c_1 / c_2$. Furthermore, we define
\begin{equation}
    \omega \coloneqq [0.0,0.25] \cup [0.75,1.0], \quad \text{and} \quad \Omega_R \coloneqq \Omega \setminus \omega = [0.25,0.75].
\end{equation}
Note that in one space dimension, the requirement \eqref{eq:Tcondition} on $T$ already implies the GCC. Let $d(x,y) \coloneqq \vert x - y \vert$, $x, y \in \Omega$, be the Euclidean metric. Then, we set
\begin{equation}
    d_c(x,y) \coloneqq \mathbbm{1}_{\Omega_1} c_1(x)^{-1} d(x,y) + \mathbbm{1}_{\Omega_2} d(x,y), \quad x,y \in \Omega.
\end{equation} 
Thus, for $\omega$ as above, $c_2 = 1$ fixed and $c_1 > 1.0$, the condition \eqref{eq:Tcondition} translates to $T > 0.25(1+c_1^{-1})$.

\noindent Figure \ref{fig:jumpCoefs:contrast:2.5} shows that for $T$ large enough, the errors $\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega_R))}$ and $\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega_R))}$ converge with order $\mathcal{O}(h^k)$. In contrast, if $T$ is chosen too small such that the GCC is not satisfied, the errors do not converge optimally. 
Figures \ref{fig:jumpCoefs:quality} and \ref{fig:jumpCoefs:increasingcontrast} show that ...

@@

\noindent As a second example, we consider the case where the data is only given in one half of the interval, i.e. $\omega = [0,0.3]$. The GCC is fulfilled if 
\begin{equation}
    T > 2 \left[ (p_2 - p_1) + \frac{p_1 - 0.3 + 1 - p_2}{c_1} \right]. 
\end{equation}
Figure \ref{fig:jumpCoefs:multipleJumps:noGCC} shows once more that the approximate solution matches the exact solution well if the GCC is fulfilled, but fails to do so if the GCC is not fulfilled.

@@

\subsubsection{Simple example} 
As a simple test case, we can extend the one-dimensional exact solution defined in \eqref{eq:1D:exact:simple} to the two-dimensional case by setting $u(x,y;t) = u(x;t)$, i.e. we consider the original solution in the $x$-direction. We define the data domain $\omega = \Omega \setminus [0.25,0.75]^2$ and note that $\omega$ fulfills the GCC. To ensure that the data domain and the interface $\Gamma$ are meshed exactly, we consider a quadrilateral mesh with $2^{L+1}$, $L \in \{1,2,3,4\}$, elements in each direction. We consider the final time $T = 0.75$  and set the time step size to $\Delta t = (1/2)^{L+1}$. As before, we fix $c_2 = 1$ and consider different values for $c_1$. The results of the experiment are presented in Fig. \ref{fig:2DjumpCoefs}.

@@

            \end{groupplot}
        \end{tikzpicture}
    \end{center}
    \caption{For $c_1 = 2.5$, $\Vert \dt(u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega))}$ (solid) and $\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega))}$ (dashed) converge quasi-optimally with order $\mathcal{O}(h^k)$ for polynomial degrees $k \in \{2,3\}$ (left) and the ratio between the approximation error and the $L^2$-bestapproximation error (right).}
    \label{fig:2DjumpCoefs}
  \end{figure}

@@

~\jp{I am not sure whether this section really fits into the narrative ... Shall we just remove it?}
We consider the same geometrical setup as in Sec. \ref{sec:numex:1D:simple} with a more intricate reference solution considered in \cite{BDE22,MEMM19}. With $u_0 (x) := \frac{1}{100} \exp(-(20(x-\frac{1}{5}))^2)$, we consider the exact solution 
\begin{equation}\label{eq:1D:complicatedSol}
    u(x,t) = \begin{cases}
      \sum_{k \ge 0} \left( \frac{c_2-c_1}{c_2+c_1}\right)^k \left( u_0 (k + x - c_1 t) - u_0(k-x-c_1 t)\right), &x \in \Omega_1, \\
      \left( \frac{2c_1}{c_2+c_1}\right) \sum_{k \ge 0} \left( \frac{c_2 - c_1}{c_2 + c_1}\right)^k u_0 \left( \frac{c_1}{c_2} \left( x - \frac{1}{2}\right) +k + \frac{1}{2} - c_1 t \right), & x \in \Omega_2. 
    \end{cases}
\end{equation}
This solution describes a wave traveling towards the interface $\Gamma$. Due to the jump in the wave speed, a part of the wave is reflected from the interface into the domain $\Omega_1$, while the other part passes through the interface into $\Omega_2$ and oscillates stronger.

@@

  \begin{figure}[!htbp]
    \begin{center}
        \begin{tikzpicture}[scale=0.72]
            \begin{groupplot}[%
                group style={%
                group size=2 by 1,
                horizontal sep=1.5cm,
                vertical sep=2cm,
                },
            ymajorgrids=true,
            grid style=dashed,
            %ymin = 1e-3, ymax = 0.5e1,
            ]    
                \nextgroupplot[width=9cm,height=7cm,domain=0:4,xmode=linear,ymode=log, xlabel={}, ylabel={}, title={$L^2$-best approximation ($k = 2$)}, %cycle list name=paulcolors, 
                legend pos=south west, 
                %yticklabels={,,}, 
                xlabel={$L$}, ylabel={$\Vert (u-L_{\Delta t} \ul_1) \Vert_{L^\infty(0,T;L^2(\Omega))}$},
                ]
                \addplot+[discard if not={order}{2},discard if not={contrast}{1.0},line width=1.1pt] table [x=L, y=bestapprox-L-infty-L2-error-u, col sep=comma] {../data/weakDisc_noRestrict_1D_jumpingCoefs_k2.csv};
                \addplot+[discard if not={order}{2},discard if not={contrast}{1.5},line width=1.1pt] table [x=L, y=bestapprox-L-infty-L2-error-u, col sep=comma] {../data/weakDisc_noRestrict_1D_jumpingCoefs_k2.csv};
                \addplot+[discard if not={order}{2},discard if not={contrast}{2.0},line width=1.1pt] table [x=L, y=bestapprox-L-infty-L2-error-u, col sep=comma] {../data/weakDisc_noRestrict_1D_jumpingCoefs_k2.csv};
                %\addplot+[discard if not={order}{2},discard if not={contrast}{2.0},line width=1.1pt,dashed] table [x=L, y=L-infty-L2-error-u, col sep=comma] {../data/weakDisc_noRestrict_1D_jumpingCoefs_k2.csv};

@@

                \legend{exact,$L=4$, $L=5$}
            \end{groupplot}
        \end{tikzpicture}
    \end{center}
    \caption{The $L^2$-bestapproximation error of \eqref{eq:1D:complicatedSol} with $k = 2$ for $c_2 = 1$ and $c_1 \in \{1.0,1.5,2.0\}$ (left) and the approximation of the time derivative of \eqref{eq:1D:complicatedSol} with $k = 3$ for $c_1 = 2.5$ for $L = 4,5$, $h = 1/2^{L+1} = \Delta t$ (right).}
    %\label{fig:2DjumpCoefs}
  \end{figure}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}[scale=3]
        \draw[-stealth] (-0.15,0) -- (1.15,0); 
        \draw[-stealth] (0,-0.15) -- (0,1.15);
        \draw (0,0) -- (1,0) -- (1,1) -- (0,1) -- cycle;
        \draw[fill=gray!40,draw=none,opacity=0.5] (0,0) -- (0.5,0) -- (0.5,1) -- (0,1) -- cycle;
        \draw[-stealth] (0.5,0) -- (1,0.5) -- (0.5,1);
        \node at (-0.15,1) {\footnotesize $T$}; 
        \node at (0.75,-0.15) {\footnotesize $\sup_{x \in \Omega} \operatorname{dist}_c(x,\omega_T)$}; 
        \node at (1.15,0.5) {\footnotesize $t_0$};
        \draw (0.7,0) to[bend right=45] (0.65,0.15); 
        \node at (0.6,0.05) {\footnotesize $\alpha$};
        \node at (0.25,0.5) {\footnotesize $\omega_T$};
        \begin{scope}[xshift=2cm]
            \draw[-stealth] (-0.15,0) -- (1.15,0); 
            \draw[-stealth] (0,-0.15) -- (0,1.15);
            \draw (0,0) -- (1,0) -- (1,1) -- (0,1) -- cycle;
            \draw[fill=gray!40,draw=none,opacity=0.5] (0,0) -- (0.25,0) -- (0.25,1) -- (0,1) -- cycle;
            \draw[fill=gray!40,draw=none,opacity=0.5] (0.75,0) -- (1,0) -- (1,1) -- (0.75,1) -- cycle;
            \draw[-stealth] (0.25,0) -- (0.75,0.5);
            \node at (-0.15,1) {\footnotesize $T$}; 
            \node at (0.5,-0.15) {\footnotesize $2 \sup_{x \in \Omega} \operatorname{dist}_c(x,\omega_T)$}; 
            \node at (0.67,0.5) {\footnotesize $t_0$};
            \draw (0.45,0) to[bend right=45] (0.4,0.15); 
            \node at (0.35,0.05) {\footnotesize $\alpha$};
            \node at (0.125,0.5) {\footnotesize $\omega_T$};
            \node at (0.875,0.5) {\footnotesize $\omega_T$};
        \end{scope}
    \end{tikzpicture}
    \caption{(\textbf{Can be deleted lateron, only for our own understanding:}) Illustration of the GCC and condition \eqref{eq:Tcondition} in one space dimension. The geometrical setup where the data is given only in one half of the domain is the worst case, because the ray can be reflected at the boundary. We still calculate that $1 = \tan(\alpha) = t_0 / \sup_{x \in \Omega} \operatorname{dist}_c(x,\omega_T)$, i.e. $T > 2 t_0 = 2\sup_{x \in \Omega} \operatorname{dist}_c(x,\omega_T)$ is required for the GCC. Note we can also assume that $c_1 = c_2 = 1$, because the metric is rescaled wrt to the wavespeed. Could probably show that the GCC and the condition on $T$ are in fact equivalent in one dimension.}
\end{figure}

>>>>

\noindent In this article, we introduce a discontinuous in time finite element method to solve the data assimilation problem \eqref{eq:waveEquation}-\eqref{eq:IF-conditions}+\eqref{eq:dataMatch} in the heterogeneous case. In particular, we want to investigate numerically if the following assumption, motivated by Thm. \ref{thm:Lipschitz} for the homogeneous case, is justified.

@@

That is, we would like to understand whether the Lipschitz stability property is preserved in presence of strong heterogeneity in the medium if the GCC holds. This question is of high practical importance since discontinuous media appear in many applications, e.g.\ in seismology. If the corresponding data assimilation problems were indeed subject to the poor logarithmic stability property ensured by \eqref{eq:FilippasEstimateLog}, this would pose an enormous challenge for the design of accurate and realiable numerical methods for their resolution.

@@

\section{Numerical experiments}
\noindent In this section, we present numerical experiments carried out with the proposed method. The examples are implemented using the \texttt{FEniCSx} library \cite{BarattaEtal2023,BasixJoss,ScroggsEtal2022,AlnaesEtal2014}. 
~\jp{We need to prepare some reproduction material ...}
\subsection{Examples in one space dimension}
\noindent We consider the unit interval $\Omega \coloneqq (0,1) \subset \mathbb{R}$. For a refinement level $L \in \mathbb{N}$, we partition the domain into $2^{L+1}$ uniform elements such that $h = 1/2^{L+1}$. The time step is chosen as $\Delta t = h$.

\subsubsection{Simple Example}\label{sec:numex:1D:simple}
\noindent We define the subdomains $\Omega_1 = (0,0.5)$ and $\Omega_2 = (0.5,1.0)$ such that $\bar{\Omega} = \overline{\Omega}_1 \cup \overline{\Omega}_2$. Then, we fix the wavespeed $c_2 := c \vert_{\Omega_2} = 1$ and consider different values for $c_1 := c \vert_{\Omega_1}$. As a first example, we consider the following exact solution \cite{MHI08}
\begin{equation}\label{eq:1D:exact:simple}
    u(x,t) := \begin{cases}
        \cos(w_1 c_1 t) \cos(w_1(x-0.5)), & x \in \Omega_1, \\
        \cos(w_2 c_2 t) \cos(w_2(x-0.5)), & x \in \Omega_2,
    \end{cases}
\end{equation}
where we set $w_1 = 3 \pi$ and $w_2 = w_1 c_1 / c_2$. Furthermore, we define
\begin{equation}
	\omega \coloneqq [0.0,0.25] \cup [0.75,1.0], \quad \text{and} \quad \Omega_R \coloneqq \bar{\Omega} \setminus \omega = [0.25,0.75].
\end{equation}
Note that in one space dimension, the requirement \eqref{eq:Tcondition} on $T$ already implies the GCC. Let $d(x,y) \coloneqq \vert x - y \vert$, $x, y \in \Omega$, be the Euclidean metric. Then, we set
\begin{equation}
    d_c(x,y) \coloneqq \mathbbm{1}_{\Omega_1} c_1(x)^{-1} d(x,y) + \mathbbm{1}_{\Omega_2} d(x,y), \quad x,y \in \Omega.
\end{equation} 
Thus, for $\omega$ as above, $c_2 = 1$ fixed and $c_1 > 1.0$, the condition \eqref{eq:Tcondition} translates to $T > 0.25(1+c_1^{-1})$.

\noindent Figure \ref{fig:jumpCoefs:contrast:2.5} shows that for $T$ large enough, the errors $\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega_R))}$ and $\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega_R))}$ converge with order $\mathcal{O}(h^k)$. In contrast, if $T$ is chosen too small such that the GCC is not satisfied, the errors do not converge optimally.

In Fig.~\ref{fig:jumpCoefs:increasingcontrast} the dependence of the errors for increasing $c_1$ and 
leaving $c_2 = 1$ fixed (i.e.\ increasing the contrast value) is investigated. When $T$ is chosen such that the GCC is barely fulfilled we observe that the errors grow polynomially in the contrast, more precisely like $\mathcal{O}(c_1^3)$.
This is an important observation, since an exponential growth in the contrast would have rendered the corresponding problems nearly impossible to resolve numerically.
Still, compared to the best approximation error which grows quadratically we 
loose a factor of $c_1$. This entails that a strong refinement of the discretization is required to maintain the accuracy of the numerical solution as the contrast increases. We illustrate this problem in Fig.~\ref{fig:jumpCoefs:quality}. For $c_1 = 2.5$ the exact solution is already well-captured by the numerical solution on refinement level $L=2$. However, for $c_1 = 5.5$ this refinement level is clearly insufficient.

@@

\noindent As a second example, we consider the case where the data is only given in one half of the interval, i.e. $\omega = [0,0.3]$. The GCC is fulfilled if 
\begin{equation}
    T > 2 \left[ (p_2 - p_1) + \frac{p_1 - 0.3 + 1 - p_2}{c_1} \right]. 
\end{equation}
Fig.~\ref{fig:jumpCoefs:multipleJumps:noGCC} shows once more that the approximate solution matches the exact solution well if the GCC is fulfilled, but fails to do so if the GCC is not fulfilled.

@@

\subsubsection{Simple example} 
As a simple test case, we can extend the one-dimensional exact solution defined in \eqref{eq:1D:exact:simple} to the two-dimensional case by setting $u(x,y;t) = u(x;t)$, i.e. we consider the original solution in the $x$-direction. We define the data domain $\omega = \Omega \setminus [0.25,0.75]^2$ and note that $\omega$ fulfills the GCC. To ensure that the data domain and the interface $\Gamma$ are meshed exactly, we consider a quadrilateral mesh with $2^{L+1}$, $L \in \{1,2,3,4\}$, elements in each direction. We consider the final time $T = 0.75$  and set the time step size to $\Delta t = (1/2)^{L+1}$. As before, we fix $c_2 = 1$ and consider different values for $c_1$. The results of the experiment are presented in Fig. \ref{fig:2DjumpCoefs}. 
For $c_1 = 2.5$, $\Vert \dt(u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega))}$ and $\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega))}$ converge quasi-optimally with order $\mathcal{O}(h^k)$ for polynomial degrees $k \in \{2,3\}$. Similar as already observed in the one-dimensional case the right plot in Fig.~\ref{fig:2DjumpCoefs} shows that these errors are more sensitive to inreasing $c_1$ than the corresponding best approximation errors. 
In particular, the ratio for $\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega))}$ with $k=2$ appears to grow as $\mathcal{O}(c_1^{3/2})$ which renders it very demanding to maintain a resonable accuracy as the contrast increases.

@@

            \end{groupplot}
        \end{tikzpicture}
    \end{center}
    \caption{Left: Convergence of $\Vert \dt(u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega))}$ (solid) and $\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega))}$ (dashed) for $c_1 = 2.5$. Right: Ratio between these errors and the $L^2$-bestapproximation error for increasing $c_1$.}
    \label{fig:2DjumpCoefs}
  \end{figure}

@@

Let us consider a final two-dimensional example to illustrate the high computational demand 
required to simulate wave propagation in heterogeneous media. 
We consider the same geometrical setup as in Sec. \ref{sec:numex:1D:simple} with a more intricate reference solution considered in \cite{BDE22,MEMM19}. With $u_0 (x) := \frac{1}{100} \exp(-(20(x-\frac{1}{5}))^2)$, we consider the exact solution 
\begin{equation}\label{eq:1D:complicatedSol}
    u(x,t) = \begin{cases}
      \sum_{k \ge 0} \left( \frac{c_2-c_1}{c_2+c_1}\right)^k \left( u_0 (k + x - c_1 t) - u_0(k-x-c_1 t)\right), &x \in \Omega_1, \\
      \left( \frac{2c_1}{c_2+c_1}\right) \sum_{k \ge 0} \left( \frac{c_2 - c_1}{c_2 + c_1}\right)^k u_0 \left( \frac{c_1}{c_2} \left( x - \frac{1}{2}\right) +k + \frac{1}{2} - c_1 t \right), & x \in \Omega_2. 
    \end{cases}
\end{equation}
This solution describes a wave traveling towards the interface $\Gamma$. Due to the jump in the wave speed, a part of the wave is reflected from the interface into the domain $\Omega_1$, while the other part passes through the interface into $\Omega_2$ and oscillates stronger.

In Fig.~\ref{fig:2D-complicated-sol} we consider the approximation of this function by the $L^2$-interpolation. The plot on the right shows that for $c_1 = 2.5$ the velocity is not captured well yet even for $k=3$ on the finest refinement level we considered. In terms of convergence rates, we merely observe $\mathcal{O}(h^2)$ for $k=2$. In view of the previous experiments it is clear that this poses a serious obstruction to obtaining accurate solutions to the corresponding unique continuation problem.

@@

  \begin{figure}[!htbp]
    \begin{center}
        \begin{tikzpicture}[scale=0.72]
            \begin{groupplot}[%
                group style={%
                group size=2 by 1,
                horizontal sep=1.5cm,
                vertical sep=2cm,
                },
            ymajorgrids=true,
            grid style=dashed,
            %ymin = 1e-3, ymax = 0.5e1,
            ]    
                \nextgroupplot[width=9cm,height=7cm,domain=0:4,xmode=linear,ymode=log, xlabel={}, ylabel={}, title={$L^2$-best approximation ($k = 2$)}, %cycle list name=paulcolors, 
                legend pos=south west, 
                %yticklabels={,,}, 
                xlabel={$L$}, 
		%ylabel={$\Vert (u-L_{\Delta t} \ul_1) \Vert_{L^\infty(0,T;L^2(\Omega))}$},
                ]
                \addplot+[discard if not={order}{2},discard if not={contrast}{1.0},line width=1.1pt] table [x=L, y=bestapprox-L-infty-L2-error-u, col sep=comma] {../data/weakDisc_noRestrict_1D_jumpingCoefs_k2.csv};
                \addplot+[discard if not={order}{2},discard if not={contrast}{1.5},line width=1.1pt] table [x=L, y=bestapprox-L-infty-L2-error-u, col sep=comma] {../data/weakDisc_noRestrict_1D_jumpingCoefs_k2.csv};
                \addplot+[discard if not={order}{2},discard if not={contrast}{2.0},line width=1.1pt] table [x=L, y=bestapprox-L-infty-L2-error-u, col sep=comma] {../data/weakDisc_noRestrict_1D_jumpingCoefs_k2.csv};
                %\addplot+[discard if not={order}{2},discard if not={contrast}{2.0},line width=1.1pt,dashed] table [x=L, y=L-infty-L2-error-u, col sep=comma] {../data/weakDisc_noRestrict_1D_jumpingCoefs_k2.csv};

@@

                \legend{exact,$L=4$, $L=5$}
            \end{groupplot}
        \end{tikzpicture}
    \end{center}
    \caption{The $L^2$-bestapproximation error of \eqref{eq:1D:complicatedSol} with $k = 2$ for $c_2 = 1$ and $c_1 \in \{1.0,1.5,2.0\}$ (left) and the approximation of the time derivative of \eqref{eq:1D:complicatedSol} with $k = 3$ for $c_1 = 2.5$ for $L = 4,5$, $h = 1/2^{L+1} = \Delta t$ (right).}
    \label{fig:2D-complicated-sol}
  \end{figure}

@@

~\jp{Remove Fig.~\ref{fig:GCC-illustration} ?}

\begin{figure}
    \centering
    \begin{tikzpicture}[scale=3]
        \draw[-stealth] (-0.15,0) -- (1.15,0); 
        \draw[-stealth] (0,-0.15) -- (0,1.15);
        \draw (0,0) -- (1,0) -- (1,1) -- (0,1) -- cycle;
        \draw[fill=gray!40,draw=none,opacity=0.5] (0,0) -- (0.5,0) -- (0.5,1) -- (0,1) -- cycle;
        \draw[-stealth] (0.5,0) -- (1,0.5) -- (0.5,1);
        \node at (-0.15,1) {\footnotesize $T$}; 
        \node at (0.75,-0.15) {\footnotesize $\sup_{x \in \Omega} \operatorname{dist}_c(x,\omega_T)$}; 
        \node at (1.15,0.5) {\footnotesize $t_0$};
        \draw (0.7,0) to[bend right=45] (0.65,0.15); 
        \node at (0.6,0.05) {\footnotesize $\alpha$};
        \node at (0.25,0.5) {\footnotesize $\omega_T$};
        \begin{scope}[xshift=2cm]
            \draw[-stealth] (-0.15,0) -- (1.15,0); 
            \draw[-stealth] (0,-0.15) -- (0,1.15);
            \draw (0,0) -- (1,0) -- (1,1) -- (0,1) -- cycle;
            \draw[fill=gray!40,draw=none,opacity=0.5] (0,0) -- (0.25,0) -- (0.25,1) -- (0,1) -- cycle;
            \draw[fill=gray!40,draw=none,opacity=0.5] (0.75,0) -- (1,0) -- (1,1) -- (0.75,1) -- cycle;
            \draw[-stealth] (0.25,0) -- (0.75,0.5);
            \node at (-0.15,1) {\footnotesize $T$}; 
            \node at (0.5,-0.15) {\footnotesize $2 \sup_{x \in \Omega} \operatorname{dist}_c(x,\omega_T)$}; 
            \node at (0.67,0.5) {\footnotesize $t_0$};
            \draw (0.45,0) to[bend right=45] (0.4,0.15); 
            \node at (0.35,0.05) {\footnotesize $\alpha$};
            \node at (0.125,0.5) {\footnotesize $\omega_T$};
            \node at (0.875,0.5) {\footnotesize $\omega_T$};
        \end{scope}
    \end{tikzpicture}
    \caption{(\textbf{Can be deleted lateron, only for our own understanding:}) Illustration of the GCC and condition \eqref{eq:Tcondition} in one space dimension. The geometrical setup where the data is given only in one half of the domain is the worst case, because the ray can be reflected at the boundary. We still calculate that $1 = \tan(\alpha) = t_0 / \sup_{x \in \Omega} \operatorname{dist}_c(x,\omega_T)$, i.e. $T > 2 t_0 = 2\sup_{x \in \Omega} \operatorname{dist}_c(x,\omega_T)$ is required for the GCC. Note we can also assume that $c_1 = c_2 = 1$, because the metric is rescaled wrt to the wavespeed. Could probably show that the GCC and the condition on $T$ are in fact equivalent in one dimension.}
\label{fig:GCC-illustration}
\end{figure}
